---
title: "RFQAmodel Package Development"
author: "Clare E. West"
date: "09/07/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

### RFQAmodel package

Today I am going to learn how to make a package.

This package will:

* read in at process the features
* get the features in a flexible way
* have a function to label models as correct/incorrect
* plot the ROC and ROC-like curves we use, flexibly
* have some kind of test

### Resources
* [R Packages](http://r-pkgs.had.co.nz/intro.html), Hadley Wickham 

### Dependencies
* devtools
* roxygen2
* testthat

```{r, eval=FALSE}
install.packages(c("devtools", "roxygen2", "testthat"))
devtools::install_github("r-lib/devtools")
```

```{r}
library(devtools)
library(roxygen2)
library(testthat)
```

### Elements of a package
* [ ] R code: R/ directory where the code lives 
* [x] Package metadata: DESCRIPTION describes the package, license and contact information
* [ ] Documentation: documentation to describe each function; created using roxygen2
* [ ] Vignettes: long-form documents that show how to combine parts of the package to solve real problems; created using R markdown
* [ ] Tests: unit tests that define correct behaviour and alert you when functions break; uses testthat
* [ ] Namespace: NAMESPACE file defines functions made available to other packages and requirements from other packages; created using roxygen2
* [ ] External data: data/ directory (e.g. examples)
* [ ] Compiled code: src/ directory to include compiled C/C++ code
* [ ] Other components: including demo/ exec/ po/ and tools/
* [ ] Git and github (via Rstudio)
* [ ] Automated checking
* [ ] Release


#### Create the package

I found that lots of the devtools functions mentioned in the R Packages book are now part of the usethis package since a [big update to devtools](https://www.tidyverse.org/articles/2018/10/devtools-2-0-0/)

```{r}
usethis::create_package("~/clareR/RFQAmodelr")
```

This will create RFQAmodelr, DESCRIPTION and NAMESPACE file.

DESCRIPTION FILE:

```
Package: RFQAmodelr
Title: What the Package Does (One Line, Title Case)
Version: 0.0.0.9000
Authors@R:
    person(given = "First",
           family = "Last",
           role = c("aut", "cre"),
           email = "first.last@example.com",
           comment = c(ORCID = "YOUR-ORCID-ID"))
Description: What the package does (one paragraph).
License: What license it uses
Encoding: UTF-8
LazyData: true
```

DESCRIPTION file can be edited manually or via more usethis functions.

For example, to add required packages:

```{r}
usethis::use_package("dplyr")
#> Setting active project to '/homes/west/clareR/RFQAmodelr'
#> Adding 'dplyr' to Imports field in DESCRIPTION
#> Refer to functions with `dplyr::fun()`

```
Other useful things:

```{r}
options(usethis.full_name = "Clare E. West")
usethis::use_mit_license()

use_readme_md()
#> Writing 'README.md'
#> Modify 'README.md'
```

### R code:

Create an `.R` file and R/ directory (if it does not already exist)

```{r}
use_r("hello")
#>Creating 'R/'
#> Modify 'R/hello.R'

## Add a function into hello.R
say_hello <- function(name="stranger"){
  print(paste("Hello, ", name, "!", sep=""))
}

## Load the package
devtools::load_all()
#>Loading RFQAmodelr

RFQAmodelr::say_hello()
#> [1] "Hello, stranger!"

RFQAmodelr::say_hello("Clare")
#> [1] "Hello, Clare!"

```

It's not possible to have subdirectories, so if files need to be grouped use a common prefix: `prefix-*.R`

Packages are loaded in two steps: when it is built (e.g. by CRAN) all the code in R/ is executed and the results are saved; when it is loaded (library() or require()), the cached results are made available. 

Don't run code at the top-level of a package, only create objects (mostly functions)

Be careful using functions that depend on global options (which may differ by user) e.g. read.csv() uses the option stringsAsFactors

To display a message when the package loads, use `.onAttach()`

```{r, eval=FALSE}
.onAttach <- function(libname, pkgname) {
  packageStartupMessage("Welcome to my package")
}
```


### Object documentation

Object documentation is accessed by `?` or `help()` and stored in `.Rd` files in the `man/` directory. The syntax is loosely based on LaTeX.

roxygen2 turns formatted comments into `.Rd` documentation files . roxygen comments start with `#'`

Documentation workflow:

* add roxygen comments to the `.R` files
* Run `devtools::document()` (or Cmd + shift + D in RStudio) to convert roxygen comments to `.Rd` files
* Preview documentation with `?`
* Repeat



```{r}
#' Say hello
#'
#' Receive a greeting from RFQAmodelr
#'
#' @param name string. Your name
#' @return Returns a greeting, personalised using \code{name} if provided.
#' @examples
#' say_hello()
#' say_hello("Clare")
say_hello <- function(name="stranger"){
  message = paste("Hello, ", name, "!", sep="")
  return(message)
}

```

Components: 

* The first sentence is the title of the documentation (use sentence case, but don't use a full stop)
* The second paragraph is the description and should briefly describe what the function does
* The optional third and further paragraphs are included in the details, shown after the argument description. 
* Blocks are broken up into tags: `@tag details`
* Most functions will have the tags `@param`, `@examples`, `@return`
* Examples must run without errors as they are run automatically, but `\dontrun{}` will prevent this (e.g. to demonstrate code that causes an error). They can be stored in separate files and use `@example path/relative/to/package/root`
* Arbitrary sections can be used: `@section My Section:`
* Params can be inherited from other function documentation using `@inheritParams source_function`

### Vignettes

A vignette is a guide to using the package, and each package may have multiple vignettes. They can be made using the R markdown vignette engine provided by knitr. 

```{r}
usethis::use_vignette("RFQAmodelr-guide")
#> Adding 'knitr' to Suggests field in DESCRIPTION
#> Setting VignetteBuilder field in DESCRIPTION to 'knitr'
#> Adding 'inst/doc' to '.gitignore'
#> Creating 'vignettes/'
#> Adding '*.html', '*.R' to 'vignettes/.gitignore'
#> Adding 'rmarkdown' to Suggests field in DESCRIPTION
#> Writing 'vignettes/RFQAmodelr-guide.Rmd'
#> Modify 'vignettes/RFQAmodelr-guide.Rmd'
```

This generates the template for a vingette, which uses R markdown. 


### Testing

Implement formal automated (AKA unit) testing using the testthat package, rather than informal testing through the command line. 

```{r}
usethis::use_testthat()
#> Adding 'testthat' to Suggests field in DESCRIPTION
#> Creating 'tests/testthat/'
#> ✔ Writing 'tests/testthat.R'
#> Call `use_test()` to initialize a basic test file and open it for editing.
```





